\subsection{MinHash}
\label{ssec:minhash}

MinHashing is an often used technique to compare big datasets. It approximates the Jaccard Coefficient, which was calculated in the previous step. In the field text mining, \citeauthor{minhash} first applied it to compare a large corpus of crawled web pages \cite{minhash}. It works by reducing the dimensionality of sequences (in the context of web pages often called documents). The algorithm assigns a signature to each document. The goal is that similar documents share many components of  their signatures. Comparing two signatures approximates the Jaccard Coefficient and can be done in a fraction of time compared to calculating the actual Jaccard Coefficient. Also, the field bio informatics deals wich large numbers of sequences. Making them quickly comparable is therefore highly beneficial. \\

In order to generate the signature of a document, it first needs to be stored in a more compact form.

\subsubsection{Matrix representation}

Let the union of the shingled documents $ S_k(A) $ and $ S_k(B) $ represent the rows of a matrix and the documents the columns. An element of the matrix is $ 1 $ if the shingle occurs in the document otherwise the element is $ 0 $. This typically leads to a sparse matrix, where most of the elements in the matrix are $ 0 $. However, this example shows a matrix which is not sparse. The data from the previous example represents the matrix $ M $. Also a matrix $ P $ of random permutations of row indices needs to be generated.\\

\begin{equation}
    \begin{split}
        P = 
        \begin{pmatrix}
            0 & 3 & 5 & 4 & 4 \\
            1 & 6 & 4 & 1 & 0 \\
            2 & 2 & 1 & 2 & 3 \\
            3 & 5 & 3 & 6 & 1 \\
            4 & 1 & 2 & 3 & 6 \\
            5 & 0 & 0 & 5 & 2 \\
            6 & 4 & 6 & 0 & 5
        \end{pmatrix}
        M = 
        \begin{pmatrix}
            1 & 0 \\
            1 & 1 \\
            1 & 1 \\
            1 & 1 \\
            1 & 0 \\
            0 & 1 \\
            0 & 1
        \end{pmatrix}   
    \end{split}
\end{equation} \\

The permutations serve in the next step to calculate the signature of the columns.\\

\subsubsection{Generation of a signature}

A signature represents a document. In the field of bio informatics, literature talks about sketches representing sequences. This is effectively the same, described with different words in different contexts.\\

With the permutations $ P $ generated before, it is possible to calculate a signature for each column.\\

To generate the signature matrix $ S $ of the columns, $ M $ is ordered by the first random permutations in $ P $. For each column in $ M $, the first row index in which the element is $ 1 $ is taken for the new signature matrix $ S $. This process is repeated for all permutations in $ P $.  It creates the matrix $ S $ with the same number of columns as $ M $ and the same number of rows as the number of columns of $ P $. Each column in $ S $ now represents a MinHash value for a document.\\

\begin{equation}
    S = 
    \begin{pmatrix}
        0 & 1 \\
        1 & 0 \\
        1 & 1 \\
        1 & 1 \\
        1 & 1
    \end{pmatrix}
\end{equation} \\

Counting the overlapping components of the signature matrix and dividing it by the total number of components, results in a value which estimates the Jaccard Coefficient. For this example the estimated Jaccard Coefficient is $ \frac{2}{5} = 0.4 $ which is fairly close to the real one with a value of $ \frac{3}{7} \approx 0.43 $.\\

The problem with this approach is that generating and storing the permutations $ P $ can take a lot of space in memory for large datasets. Also sorting the complete matrix for each permutations is a time consuming task. These problems can be overcome by using hash functions instead of permutations. The next section describes how hash functions work. Afterwards, the author explains how they can be used in the process of permutation generation.\\